# Activity File for Portland, OR

```{r chunk1, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = TRUE, message = FALSE, fig.width = 8}
library("tidyverse")
library("devtools")
#install_github("ITHIM/ITHIM", ref = "ITHIM-R")
install("~/ITHIM/")
library("ITHIM")
```

Update _trippub_ to be the filename with full path.

```{r chunk2, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE, fig.width = 8}
trippub <- "~/NHTS/data/trippub.csv"
```

Now we read in the data and subset on _HHCBSA == 38900_ (the Portland
Metropolitan area) and create a binary age category split at age 18
years.

```{r chunk3, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE, fig.width = 8}
personaltrips <- read.csv2(trippub, header = TRUE, sep = ",")
portlanddata <- personaltrips[which (personaltrips$HH_CBSA == 38900),]
portlanddata$agegroup <- ifelse(portlanddata$R_AGE <= 18, 0, 1)
```

In order to weight cycling, walking and transit durations differently
we need to separate out the respective durations.  I have included
some code to help start.

```{r chunksgy, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE, fig.width = 8}

portlanddata %>%
    group_by(TRPTRANS, HOUSEID, PERSONID, agegroup) %>%
    summarise( T = sum(TRVLCMIN) ) %>%
    spread(TRPTRANS, T, fill = 0) %>%
    ungroup() %>%
    filter( (`1` > 0) | (`2` > 0) | (`11` > 0) | (`16` > 0)) %>%
    select( c(HOUSEID, PERSONID,agegroup,`1`, `2`, `11`, `16`)) %>%
    mutate(TA = 7*(3*`1`/60 + 6*`2`/60))
```

Note that i excluded mode 15 because it did not appear in the Portland
data.  It would be much better if we renamed the modes.  We can
recycle this code I wrote a few months ago.  In it I simplified the
bus and subway modes.

```{r chunksgy2, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE, fig.width = 8}

portlanddata <- personaltrips[which(personaltrips$HH_CBSA == 38900),]

portlanddata <- with(portlanddata,{
    data.frame(HOUSEID = HOUSEID,
               PERSONID = PERSONID,
               TDCASEID = TDCASEID,
               TRPTRANS = factor(ifelse(TRPTRANS %in% c(-9,-8,-7,97), NA, TRPTRANS), levels = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20), labels = c("walk","cycle","car","SUV","van","pickup truck","golf cart/Segway","motorcycle/moped","RV","school bus","bus","paratransit/Dial-a-ride","private/charter/tour/shuttle bus","city-to-city bus (Greyhound, Megabus)","Amtrak/commuter rail","subway","taxi/limo/Uber/Lyft", "rental car/Zipcar/Car2Go", "airplane","boat/ferry/water taxi")),
               TRVLCMIN = as.numeric(TRVLCMIN),
               URBRUR  = as.factor(URBRUR),
               MSACAT = as.factor(MSACAT),
               MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
               TRPMILES = as.numeric(TRPMILES),
               agegroup = ifelse(R_AGE <= 18, "minor", "adult")
)})

portlanddata <- with(portlanddata,{
    data.frame(houseID=HOUSEID, subjectID=PERSONID, duration = TRVLCMIN, mode = TRPTRANS, agegroup = agegroup)
})
```

```{r chunksgy3, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE, fig.width = 8}

TA.df <- portlanddata %>%
    group_by(mode, houseID, subjectID, agegroup) %>%
    summarise( T = sum(duration) ) %>%
    spread(mode, T, fill = 0) %>%
    ungroup() %>%
    filter( (walk > 0) | (cycle > 0) | (subway > 0) | (bus > 0)) %>%
    select( c(houseID, subjectID, agegroup, walk, cycle, bus, subway)) %>%
    mutate(TA = 7*(3*walk/60 + 6*cycle/60 + ifelse(bus > 0, 3*15/60,0) + ifelse(subway > 0, 3*15/60,0)))

hist(TA.df$TA)
hist(log(TA.df$TA))
```

Recall that this is the distribution of travel activity for active
travelers. (TA > 0).

```{r chunksgy4, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE, fig.width = 8}
TA.summary <- TA.df %>% group_by(agegroup) %>% summarise(mean = mean(TA), sd = sd(TA))
```

```{r chunksgy5, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE, fig.width = 8}
p0.df <- portlanddata %>%
    group_by(mode, houseID, subjectID, agegroup) %>%
    summarise( T = sum(duration) ) %>%
    spread(mode, T, fill = 0) %>%
    ungroup() %>%
    group_by(agegroup) %>%
    summarise( n1 = sum((walk > 0) | (cycle > 0) | (subway > 0) | (bus > 0)), n0 = sum((walk == 0) & (cycle == 0) & (subway == 0) & (bus == 0)), n = n()) %>%
    ungroup() %>%
    mutate(p0 = n0/n) %>%
    select(agegroup, p0)
```

```{r chunksgy6, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE, fig.width = 8}
activityFile.df <- full_join(TA.summary,p0.df, by = "agegroup")
colnames(activityFile.df) <- c("ID","mean","sd","p0")
write.csv(activityFile.df, file = "./portland.baseline.csv", row.names = FALSE, quote = FALSE)
```

```{r chunksgy7, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE, fig.width = 8}
activityFile.scenario.df <- within(activityFile.df, {
    p0 <- ifelse( p0 - 0.10 >= 0, p0 - 0.10, 0)
})
write.csv(activityFile.scenario.df, file = "./portland.scenario.csv", row.names = FALSE, quote = FALSE)
```

```{r chunksgy8, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE, fig.width = 8}
CRA("portland.baseline.csv","portland.scenario.csv")
```

```{r chunksgy9, eval = TRUE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE, fig.width = 8}
activityFile.scenario.df <- within(activityFile.df, {
    p0 <- ifelse( p0 + 0.10 <= 1, p0 + 0.10, 1)
})
write.csv(activityFile.scenario.df, file = "./portland.scenario.csv", row.names = FALSE, quote = FALSE)
CRA("portland.baseline.csv","portland.scenario.csv")
```

## Archive (Unevaluated code)

In the following code chuk I changed the syntax a little.  Not sure it
was working properly before.

```{r chunk4, eval = FALSE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE, fig.width = 8}
portlanddata$transitminutes <- if_else(portlanddata$TRPTRANS %in% c(11,16), portlanddata$TRVLCMIN/2, ifelse(portlanddata$TRPTRANS %in% c(1,2), portlanddata$TRVLCMIN*1, 0))
```

In the code chunk above we have included bus and light rail trips, but only at half
the duration.  I propose we change this to a fixed amount, say the
average time it takes to walk to a bus stop, and score it as walking
for that duration.


```{r chunk5, eval = FALSE, echo = TRUE, results = "show", warning = TRUE, error = TRUE, message = TRUE, fig.width = 8}

aggregatedportland <- portlanddata %>%
  group_by(HOUSEID, PERSONID) %>%
  summarise(agegroup = first(agegroup),
            transitminutes = sum(transitminutes))


summaryData <- cbind(
  aggregate(
    aggregatedportland[, "transitminutes"],
    by = list(agegroup = aggregatedportland$agegroup),
    FUN = mean,
    na.rm = FALSE
  ),
  aggregate(
    aggregatedportland[, "transitminutes"],
    by = list(agegroup = aggregatedportland$agegroup),
    FUN = sd,
    na.rm = FALSE
  )[, 2]
)

colnames(summaryData) <- c("agegroup", "mean", "sd")

summaryData$proportion <- NA

agegroups <- unique(aggregatedportland$agegroup)

for (i in 1:length(agegroups)) {
  tempdata <- aggregatedportland[which(aggregatedportland$agegroup == agegroups[i]), ]
  summaryData[i, "proportion"] <- length(which(tempdata$transitminutes > 0))  / nrow(tempdata)
}
```
